---
title: "template for .qmd files"
format: html
toc: false
---

The goal is to create a single .qmd file which will generate the additional .qmd files which when compiled for the individual course note pages.

```{r}
#| include: false

library(tidyverse)
library(DT)
library(gt)
```


```{r}
#| echo: false
#| warning: false


metadata <- read_csv("../contents_raw.csv")

metadata <- metadata |>
  mutate(vol = str_sub(issue, 5, 6))

metadata

# create vector of filenames we will write to

outfiles <- ""

for (i in 1:nrow(metadata)) {
  outfiles[[i]] <- str_glue("v{metadata$vol[[i]]}.qmd")
}

outfiles

```

### This code chunk takes a single row from `metadata` and transposes it so we can display a table with it's entries. We also take the column names from metadata for the labels of that table. Ultimately, this is the R code that we want to write to multiple .qmd files

```{r}

output <- metadata |>
  filter(vol == "47") |>
  t()


cn <- metadata |>
  colnames()

tibble(cn, output) |>
  gt() |>
  cols_align(align = "left") |>
  tab_style(
    style= cell_text(v_align = "top"),
    locations = cells_body()
  ) |>
  tab_options( table.width = "80%", column_labels.hidden = TRUE)
```

# First, we make sure we can create a single new `.qmd` file

```{r}

new_block <- '---
title: "title"
subtitle: "subtitle"
---

## Introduction
This file was generated using R code.

\`\`\`{r}
#| echo: false
#| warning: false

library(tidyverse)
library(gt)

metadata <- read_csv("../contents_raw.csv")
metadata <- metadata |>
  mutate(vol = str_sub(issue, 5, 6))
  
output <- metadata |>
  filter(vol == "47") |>
  t()

cn <- metadata |>
  colnames()

tibble(cn, output) |>
  gt() |>
  cols_align(align = "left") |>
  tab_style(
    style= cell_text(v_align = "top"),
    locations = cells_body()
  ) |>
  tab_options( table.width = "80%", column_labels.hidden = TRUE)
\`\`\`
'

writeLines(new_block, "new_file.qmd")
```

# Experiment with str_glue

```{r}

x <- 42

test_block <- str_glue('---
title: "title #{x}"
subtitle: "subtitle"
---

\`\`\`{{r}}
\`\`\`

## Introduction
This file was generated with the secret number {x}.

')

test_block
```

# Next step is to create multiple files for each volume.

```{r}
#| include: false


# Define the content of the Quarto file

title_block <- ""

for (i in 1:nrow(metadata)) {
  title_block[[i]] <- str_glue("---
  title: \"{metadata$title[[i]]}\"
  subtitle: \"{metadata$issue[[i]]}\"
  format: html
  toc: false
  ---
  ")
}


block1 <- '
## Introduction
This file was generated using R code.

\`\`\`{r}
#| echo: false
#| warning: false

library(tidyverse)
library(gt)

metadata <- read_csv("../contents_raw.csv")
'



block3 <- 'cn <- metadata |>
  colnames()

tibble(cn, head(output, n=-1)) |>
  gt() |>
  cols_align(align = \"left\") |>
  tab_style(
    style= cell_text(v_align = \"top\"),
    locations = cells_body()
  ) |>
  tab_options( table.width = \"80%\", column_labels.hidden = TRUE)
\`\`\`
'



block2 <- ""
qmd_block <- ""

for (i in 1:nrow(metadata)) {
  block2[[i]] <- str_glue("output <- metadata |>
  mutate(v = str_sub(issue, 5, 6)) |>
  filter(v == {metadata$vol[[i]]}) |>
  t()
  
  ")
  qmd_block[[i]] <- str_c(title_block[[i]], block1, block2[[i]], block3)
}





walk2(qmd_block, outfiles, 
      \(content, path) writeLines(content, path))

```
