---
title: "Course Modules"
format: html
---

These modules have been course tested and reviewed... according to the following guidelines...

Keeping JIBLM free is a top priority. If you use JIBLM materials, we want to know! Please consider emailing our editors to let us know which notes you use, your institution, your name, and anything else you wish to share.

This table is sortable. Click on a column heading to sort by that column. Click on a title to access these notes.

```{r}
#| include: false

library(tidyverse)
library(DT)
```

```{r}
#| include: false


modules <- read_csv("modules.csv") |>
  mutate(vol = as.integer(str_sub(Issue, 5, 6)))


modules <- modules |>
  mutate( Title = str_glue("<a href= .\\M0{vol}I.pdf >{Title}</a>" ))

  
#  separate_wider_regex(Issue, patterns = c("No. ", v = "[0-9]+", " [(]", date = "[A-Za-z]+ [0-9]+", " [)]"))

```

```{r}
#| echo: false

options(DT.options = list(order = list(2, 'asc'), dom = 't'))

#materials |>
#  datatable(rownames = FALSE, style = 'bootstrap', class = "table-condensed table-striped", escape = FALSE)

modules |>
  separate_wider_regex(Issue, patterns = c("No.\\s", Volume = "[0-9]+", " [(]", date = "[A-Za-z]+\\.?\\s[0-9]+", "[)]")) |>
  mutate(Date = format(my(date), "%Y %b")) |>
  select(!date & !vol) |>
  relocate(Date, .after = Volume) |>
  rename( Issue = Volume ) |>
  datatable(rownames = FALSE, style = 'bootstrap', class = "table-condensed table-striped", escape = FALSE)
```


