---
title: "test_template"
---

this .qmd file generates additional .qmd files which when compiled will be the individual course note pages.



```{r}
#| include: false

library(tidyverse)
library(DT)
library(gt)
```




```{r}
#| echo: false
#| warning: false


metadata <- read_csv("../contents_raw.csv")

metadata <- metadata |>
  mutate(vol = str_sub(issue, 5, 6))

metadata

# create vector of filenames we will write to

outfiles <- ""

for (i in 1:nrow(metadata)) {
  outfiles[[i]] <- str_glue("v{metadata$vol[[i]]}.qmd")
}

outfiles

```


### This code chunk takes a single row from `metadata` and transposes it so we can display a table with it's entries.  We also take the column names from metadata for the labels of that table.  Ultimately, this is the R code that we want to write to multiple .qmd files

```{r}

output <- metadata |>
  filter(vol == "47") |>
  t()


cn <- metadata |>
  colnames()

tibble(cn, output) |>
  gt() |>
  cols_align(align = "left") |>
  tab_style(
    style= cell_text(v_align = "top"),
    locations = cells_body()
  ) |>
  tab_options( table.width = "80%", column_labels.hidden = TRUE)
```



### This is the code chunk that writes the R code to `new_file.qmd`

```{r}
#| include: false


# Define the content of the Quarto file
block1 <- '---
title: "My Quarto Report"
format: html
---

## Introduction
This file was generated using R code.

\`\`\`{r}
#| echo: false
#| warning: false

library(tidyverse)
library(gt)

metadata <- read_csv("../contents_raw.csv")
'



block3 <- 'cn <- metadata |>
  colnames()

tibble(cn, head(output, n=-1)) |>
  gt() |>
  cols_align(align = \"left\") |>
  tab_style(
    style= cell_text(v_align = \"top\"),
    locations = cells_body()
  ) |>
  tab_options( table.width = \"80%\", column_labels.hidden = TRUE)
\`\`\`
'



block2 <- ""
qmd_block <- ""

for (i in 1:nrow(metadata)) {
  block2[[i]] <- str_glue("output <- metadata |>
  mutate(v = str_sub(issue, 5, 6)) |>
  filter(v == {metadata$vol[[i]]}) |>
  t()
  
  ")
  qmd_block[[i]] <- str_c(block1, block2[[i]], block3)
}



# Write the content to a file named "new_flle.qmd"
writeLines(qmd_block, "new_file.qmd")

walk2(qmd_block, outfiles, 
      \(content, path) writeLines(content, path))

```


