---
title: "template for course note .qmd files"
format: html
toc: false
---

## Rendering this file will generate the .qmd files for each course notes page. Note that this will overwrite the .qmd files!

```{r}
#| include: false

library(tidyverse)
library(DT)
library(gt)
```

```{r}
#| echo: false
#| warning: false


metadata <- read_csv("contents_raw.csv")

metadata <- metadata |>
  mutate(vol = as.integer(str_sub(issue, 5, 6)))

# create vector of filenames we will write to

outfiles <- ""

for (i in 1:nrow(metadata)) {
  outfiles[[i]] <- str_glue("v{metadata$vol[[i]]}/v{metadata$vol[[i]]}.qmd")
}

outfiles

```

### This code chunk takes a single row from `metadata` and transposes it so we can display a table with it's entries. We also take the column names from metadata for the labels of that table. Ultimately, this is the R code that we want to write to multiple .qmd files

```{r}
#| echo: false


output <- metadata |>
  filter(vol == "47") |>
  select(!file1 & !file2 & !review) |>
  t()


cn <- metadata |>
  select(!file1 & !file2 & !review) |>
  colnames()

tibble(cn, output) |>
  gt() |>
  cols_align(align = "left") |>
  tab_style(
    style= cell_text(v_align = "top"),
    locations = cells_body()
  ) |>
  tab_options( table.width = "80%", column_labels.hidden = TRUE)
```

```{r}
#| include: false


# Define the content of the Quarto file


qmd_block <- ""

for (i in 1:nrow(metadata)) {

  block1 <- str_glue("---
title: \"Course Notes\"
subtitle: \"{metadata$title[[i]]}, {metadata$issue[[i]]}\"
---

\`\`\`{{r}}
#| echo: false
#| warning: false

library(tidyverse)
library(gt)

metadata <- read_csv(\"../contents_raw.csv\")

output <- metadata |>
  mutate(v = as.integer(str_sub(issue, 5, 6))) |>
  filter(v == {metadata$vol[[i]]}) |>
  select(!file1 & !file2 & !review) |>
  t()
  
cn <- metadata |>
  select(!file1 & !file2 & !review) |>
  colnames()

tibble(cn, head(output, n=-1)) |>
  gt() |>
  cols_align(align = \"left\") |>
  tab_style(
    style= cell_text(v_align = \"top\"),
    locations = cells_body()
  ) |>
  tab_options( table.width = \"80%\", column_labels.hidden = TRUE)
\`\`\`

")

  
block2 <- str_glue("    
- [Download Instructor Version](./v{metadata$vol[[i]]}I.pdf)\n

")

block3 <- str_glue("
- [Download Student Version](./v{metadata$vol[[i]]}S.pdf)\n

")

block4 <- str_glue("
- [Download Review](./v{metadata$vol[[i]]}_review.pdf)\n

")

if (!is.na(metadata$file2[[i]])) {
  block2 <- str_c(block2, block3)
}

if (metadata$review[[i]] == 1){
  block2 <- str_c(block2, block4)  
}

block5 <- str_glue("
This refereed course guide is supplied to you free of charge. In exchange, we ask only that if you reproduce a course guide, you do not sell it for more than the cost of reproduction. At JIBLM, the author retains the copyright to his or her material and many are willing to share the .tex source of their notes. If you wish to have the source in addition to the .pdf available below, we ask that you contact the author directly. The author\'s email address at the time of submission is provided above. "
)



qmd_block[[i]] <- str_c(block1, block2, block5)
  
dir.create(str_glue("v{metadata$vol[[i]]}"))
}





walk2(qmd_block, outfiles, 
      \(content, path) writeLines(content, path))

```
